---
title: "Data Wrangling Assignment"
author: "Josh Clinton & Will Doyle"
date: "9/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this lecture you are going to wrangle some exit poll data to produce a tibble for analysis!

Using the `MI_final`, create new tibble called `MI_small` containing only the following variables: `PARTYID`, `preschoice`, and `QLT20`.

Moreover, make sure the data contains only people who report voting for either Biden or Trump.

Now create two tibbles:
-  `BidenReasons` that is a count of how many Biden voters choose each reason.  
- `TrumpReasons` that is a count of how many Trump voters chose each reason.  

```{r, message=FALSE}
library(tidyverse)
load(file = "data/Final MI subset.Rdata")

MI_small <- select(MI_final, c(PARTYID, preschoice, QLT20))

MI_small <- filter(MI_small, preschoice == "Joe Biden, the Democrat" | preschoice == "Donald Trump, the Republican")
```

You can crete these tibbles in two ways.  First, using the methods we introduced in class, we can use the `count` function applied to the `filter`ed tibble to create the tibbles as follows.

```{r, message=FALSE}
BidenReasons <- count(filter(MI_small, 
                             preschoice == "Joe Biden, the Democrat"), QLT20)

TrumpReasons <-count(filter(MI_small, 
                            preschoice == "Donald Trump, the Republican"), QLT20)
```

A different way to do the same thing using pipes is as follows.  Recall that `%>%` is read as "then" so in each case we are working with `MI_small` _then_ `filter`ing to select the set of voters _then_ `count`ing `QLT20`.  (NOTE: that we will learn how we can do this in a single tibble using the `goup_by` function).

```{r, message=FALSE}
BidenReasons_1 <- MI_small %>%
                  filter(preschoice == "Joe Biden, the Democrat") %>%
                  count(QLT20)

TrumpReasons_1 <- MI_small %>%
                  filter(preschoice == "Donald Trump, the Republican") %>%
                  count(QLT20)
```


Print these to the screen and comment briefly on:

1. What are the dimensions of `MI_small`?  What do these numbers refer to? Is this how much data we have to analyze?

```{r}
dim(MI_small)

# The first number is the number of observations -- here we 1182 have Exit Poll respondents in our data tibble and the second number is the number of variables we have on each respondent - here 3.  Note that this is NOT how much data we have because some may be missing (`NA`). 
```



*1. Print the the two tibbles to the screen. How similar are the reasons that supporters of Biden and Trump give for their support? Why might that matter?*

```{r}
BidenReasons
TrumpReasons
```

What does this mean?  It seems that Trump voters are voting for a "strong leader," but  Biden voters are voting for "good judgement".  Why does this matter?  Well, if Trump voters care primarily about Trump showing strength rather than judgement, but Biden voters value judgment over strength this can have implications for how each could lose the support of his supporters.  In particular, if voters start to question Biden's judgement -- e.g., Afghanistan? -- this could bode ill for him.  In contrast, when Trump was president his voters were less responsive to many of the aspects of his presidency because his voters were voting for "strength" so long as Trump acted "strong" his voters were likely to approve.

*2. What does `NA` in these tibbles mean?*

In R, the value of `NA` refers to missing values.  It indicates if we do not have a response/data/value for that respondent on that variable.  When we have missing data in a variable we have to be careful to check what R does when it encounters them.  Some functions -- e.g., `mean` -- will not compute when there is missing data unless we tell R what to do and whether to remove missing data of not (using `na.rm=TRUE` - but that syntax is not generically applicable to every function call in R!).


*3. Now produce `tibbles` of the count without the `NA`?  (HINT: the function `is.na()` is useful for this)!*

Results:
```{r}
BidenReasons_nona <- count(filter(MI_small, 
                                  preschoice == "Joe Biden, the Democrat" & !is.na(QLT20)),
                           QLT20)
TrumpReasons_nona <-count(filter(MI_small, 
                                 preschoice == "Donald Trump, the Republican" & !is.na(QLT20)),
                          QLT20)
BidenReasons_nona 
TrumpReasons_nona 
```

Using Pipes:

```{r}
BidenReasons_nona1 <- MI_small %>%
    filter(preschoice == "Joe Biden, the Democrat" & !is.na(QLT20)) %>%
    count(QLT20)

TrumpReasons_nona1 <- MI_small %>%
    filter(preschoice == "Donald Trump, the Republican" & !is.na(QLT20)) %>%
    count(QLT20)
```

If we wanted to drop missing data caused by respondents not anwering (as opposed to respondents not being asked -- which is which we just did) we can change the `filter` to also exclude those values from `QLT20`:

```{r}
BidenReasons_nona <- count(filter(MI_small, 
          preschoice == "Joe Biden, the Democrat" & !is.na(QLT20) & QLT20 != "[DON'T READ] Don’t know/refused"), 
          QLT20)

TrumpReasons_nona <-count(filter(MI_small, preschoice == "Donald Trump, the Republican" & !is.na(QLT20) & QLT20 != "[DON'T READ] Don’t know/refused"), 
                          QLT20)

BidenReasons_nona 
TrumpReasons_nona 
```

*4. As question 2 makes, clear, we are often interested in how reasons compare between group. Using the ability to reference variables within a tibble (i.e., `$`), subtract the count of reasons given by Biden voters from the count of reasons given by Trump voters.  How meaningful is what we just did for understanding differences between Biden and Trump voters?  In particular, if you were running Election Night coverage for a major network would you present the numbers you just calculated on-air _as is_?  Why or why not?*

```{r}
TrumpReasons_nona$n - TrumpReasons_nona$n
```

Would we ever report this?  Absolutely not because the meaning of this is unclear because the difference being calculated is the difference in the number of respondents.  But this is impossible for viewers to interpret because the values are relative to the number of responses being given.  It is impossible to know from this whether it is large or small. Is 100 big?  

If you wanted to compare the support it would make much more sense to calculate the percentage of Biden voters and Trump voters who select each response and then compare the percentage point difference in support to see how much more or less likely each set of voters to select each response option.

It is also hard to interpret because all we asked for is a vector of numbers.  

```{r, warning=FALSE}
Diff <- TrumpReasons_nona$n - BidenReasons_nona$n
DiffTable <- bind_cols(BidenReasons_nona$QLT20,Diff)
colnames(DiffTable) <- c("Reasons","TrumpBidenDifference")
DiffTable
```

To get the percentage, look at the lecture notes to see how we created a count tibble with a proportion for each response category. We could create then and then subtract them across candidates. 

*Stretch Items*

*A. Create a binary variable that's set to 1 if the respondent indicated they voted for Biden.*

Lots of ways to do this.  Here is one way using pipes.

```{r}
MI_final<-MI_final%>%
  mutate(biden_binary=ifelse(PRSMI20==1,1,0))
```

Or...

```{r, eval = FALSE}
MI_final$biden_binary <- ifelse(PRSMI20==1,1,0)

MI_final <- mutate(MI_final, 
                   biden_binary=ifelse(PRSMI20==1,1,0))
```


*B. Calculate the proportion of respondents who said they voted for Biden by education level.*

```{r}
MI_final%>%
  group_by(EDUC18)%>%
  summarize(prop_biden=mean(biden_binary))
```


Comment on any patterns you see. 

*C. Now calculate the proportion of respondents who said they voted for Biden by education level AND gender. To do so we must assume that it is OK to use the responses to the variable labeled as `SEX` to measure gender.* 

```{r}
MI_final%>%
  group_by(EDUC18,SEX)%>%
  summarize(prop_biden=mean(biden_binary))
```


*Does the relationship (if any) between education level and voting for Biden differ by gender? How?*

*D. Plot the result of C.* 

```{r}
MI_final%>%
  group_by(EDUC18,SEX)%>%
  summarize(prop_biden=mean(biden_binary))%>%
  ggplot(aes(x=as.factor(EDUC18),
             y=prop_biden,
             fill=as.factor(SEX)))+
  geom_bar(position="dodge",stat="identity")+
  theme(legend.position = "none")+
  ylab("Proportion Voting for Biden")+
  xlab("Education Level")
```

Note that the x-axis is labelled using the value of `EDUC18` which is uninformative.  We will learn how to change it to something reasonable.


*E. Choose another variable and combine it with education, looking at Biden by vote share and that variable. Say WHY you chose that variable and WHAT the patterns reveal when you combine these two variables. *

Race is a powerful predictor of political opinions given the foundational and enduring impact and legacy of race on politics as well as political, social, and economic institutions.  So lets' see how the results differ by race, realizing that we may be working with a small number of cases which may make it difficult to learn much about how these relationships.

```{r}
MI_final%>%
  group_by(QRACEAI,SEX)%>%
  summarize(prop_biden=mean(biden_binary))

```


*F. Plot the result of E.*

Note the use of `recode` here along with `mutate`!


```{r, warning=FALSE}
MI_final%>%
  mutate(race_ethnicity=recode(as.character(QRACEAI),
                     "1"="White",
                     "2"="Black",
                     "3"="Hispanic/Latino",
                     "4"="Asian",
                     "5"="American Indian",
                     "6"="Other",
                     "9"="No Answer"))%>%
  mutate(Gender=recode(as.character(SEX),
                       "1"="Man",
                       "2"="Woman"))%>%
  group_by(race_ethnicity,Gender)%>%
  summarize(prop_biden=mean(biden_binary))%>%
  ggplot(aes(x=fct_reorder(race_ethnicity,prop_biden),
             y=prop_biden,
             fill=Gender))+
  geom_col(position = "dodge")+
  coord_flip()+
  ylab("Proportion Voting for Biden")+
  xlab("Race/Ethnicity")

```

