---
title: "Assignment 5"
author: "Will Doyle"
date: "11/30/2021"
output: html_document
---

```{r}
library(tidyverse)
library(tidymodels)
```

```{r}
ad<-read_rds("admit_data.rds")
```


Complete the following steps:

1. Plot yield as a function of income

```{r}
ad%>%
  mutate(income_percent=ntile(income,n=100))%>%
  group_by(income_percent)%>%
  summarize(mean_yield=mean(yield))%>%
  ggplot(aes(x=income_percent,y=mean_yield))+
  geom_point()
```


1. Calculate the proportion who yielded as a function of income decile and comment briefly on the results. 

```{r}
ad%>%
  mutate(income_percent=ntile(income,n=10))%>%
  group_by(income_percent)%>%
  summarize(mean_yield=mean(yield))
```

Higher income students are much more likely to yield than low income students-- only 20 percent of low income students attend, while all high income students attend. 

2. Run a regression with yield as a function of income and sent scores. Comment on the coefficients.


```{r}
ad_split<-initial_split(ad)

ad_train<-training(ad_split)

ad_test<-testing(ad_split)
```


```{r}
lm_fit<-linear_reg()%>%
  set_engine("lm")%>%
  set_mode("regression")
```


```{r}
yield_formula<-as.formula("yield~income+sent_scores")

yield_recipe<-recipe(yield_formula,ad_train)

```

```{r}
yield_wf<-workflow()%>%
  add_model(lm_fit)%>%
  add_recipe(yield_recipe)
```

```{r}
yield_wf<-yield_wf%>%
  fit(ad_train)
```

```{r}
yield_wf%>%tidy()
```

If a student sends their scores, the probability of yield goes up by about 10 percent. For each additional 1000 in income, the probability of attendance goes up by about 2 percent. 

3. Run a logistic regression with yield as a function of income and sent scores. Again comment on the coefficients. 

```{r}
ad<-ad%>%
  mutate(yield_f=ifelse(yield==1,"Yes","No"))%>%
  mutate(yield_f=fct_relevel(yield_f,ref="No"))
```



```{r}
ad_split<-initial_split(ad)

ad_train<-training(ad_split)

ad_test<-testing(ad_split)
```



```{r}

yield_formula<-as.formula("yield_f~income+sent_scores")

yield_recipe<-recipe(yield_formula,ad_train)
```



```{r}
logit_mod<-logistic_reg()%>%
  set_engine("glm")%>%
  set_mode("classification")
```



```{r}
yield_wf<-workflow()%>%
  add_model(logit_mod)%>%
  add_recipe(yield_recipe)
```

```{r}
yield_wf<-yield_wf%>%
  fit(ad_train)
```

```{r}
yield_wf%>%tidy()
```

Income has a positive and statistically significant relationship with yield, as does sent scores. 

4. Calculate the accuracy, specificity and sensitivity of the predictions from your logistic regression above. Comment briefly on each. 

```{r}

##  Check baseline first

ad%>%
  summarize(mean(yield))

yield_wf%>%
  predict(ad_test)%>%
  bind_cols(ad_test)%>%
  accuracy(truth=yield_f,estimate=.pred_class,event_level="second")
```

Overall, 68 percent of admitted students enroll. The model correctly classifies about 77 percent of students overall, better than what one could do simply by chance. 


```{r}
yield_wf%>%
  predict(ad_test)%>%
  bind_cols(ad_test)%>%
  sens(truth=yield_f,estimate=.pred_class,event_level="second")
```


The model correctly classifies 82 percent of those who actually do enroll.

```{r}
yield_wf%>%
  predict(ad_test)%>%
  bind_cols(ad_test)%>%
  spec(truth=yield_f,estimate=.pred_class,event_level="second")
```

The model correctly classifies 67 percent of those who do not enroll. 


5. Calculate the AUC from the logistic regression and state whether you have achieved good model fit. 

```{r}
yield_wf%>%
  predict(ad_test,type="prob")%>%
  bind_cols(ad_test)%>%
  roc_auc(truth=yield_f,estimate=.pred_Yes,event_level="second")
```

The AUC for the model is .86, indicating a reasonable degree of both sensitivity and specificity at a variety of different thresholds. 
