---
title: "GE3_ides"
author: "Will Doyle"
date: "11/4/2021"
output: html_document
---

```{r}
library(tidymodels)
library(tidyverse)
library(tidytext)
```

```{r}
fix_na<-function(x){
  ifelse(x=="N/A",NA,x)
}
```


```{r}

mv_train<-readRDS("movie_train.Rds")%>%
  mutate(across(where(is.character),fix_na))%>%
  mutate(across(.cols=c(boxoffice_a,metascore,imdbrating,imdbvotes),as.numeric))%>%
  filter(across(where(is.numeric), ~!is.na(.x)))%>%
  filter(!is.na(rated))%>%
  filter(!is.na(boxoffice_a))%>%
  mutate(log_box_office=log(boxoffice_a))

mv_test<-readRDS("movie_test.Rds")%>%
  mutate(across(where(is.character),fix_na))%>%
  mutate(across(.cols=c(metascore,imdbrating,imdbvotes),as.numeric))%>%
  filter(across(where(is.numeric), ~!is.na(.x)))%>%
  filter(!is.na(rated))

```


## Idea 1: Using words in plot summaries

There may be certain terms that are used in movie plots that are 
associated with better box office earnings. One might imagine that "exciting" would 
sell better than "epistemology." Below, I select the most commonly occurring words and create a series of dummy variables for each word. 

```{r}

## Select common words from plots in _testing_ data

mv_plot_words_test<-mv_test%>%
select(imdbid,plot)%>%
  group_by(imdbid)%>%
  unnest_tokens(input=plot,output=word,token="words")%>%
  anti_join(stop_words,by="word")%>%
  ungroup()

common_words_test<-mv_plot_words_test%>%
  group_by(word)%>%
  count()%>%
  arrange(-n)%>%
  ungroup()%>%
  filter(n>=10)%>%
  select(word)%>%
  as_vector()


mv_plot_words_train<-mv_train%>%
select(imdbid,plot)%>%
  group_by(imdbid)%>%
  unnest_tokens(input=plot,output=word,token="words")%>%
  anti_join(stop_words,by="word")%>%
  ungroup()

common_words_train<-mv_plot_words_train%>%
  group_by(word)%>%
  count()%>%
  arrange(-n)%>%
  ungroup()%>%
  filter(n>=10)%>%
  select(word)%>%
  as_vector()

common_words<-common_words_train[common_words_train%in%common_words_test]

for (word in common_words){  

word_name<-paste0(word,"_wd")
  
mv_train<-mv_train%>%
  mutate(temp =ifelse(str_detect(plot,word),1,0))
  
mv_train[word_name]<-mv_train$temp
}
  

for (word in common_words){  

word_name<-paste0(word,"_wd")
  
mv_test<-mv_test%>%
  mutate(temp =ifelse(str_detect(plot,word),1,0))
  
mv_test[word_name]<-mv_test$temp
}
```


## Actors: Who's REALLY an A-Lister?

The same could be true of actors-- are there certain actors that are a box office draw? Below
I select the most common actors and create a binary variable for their presence or absence in
the movie. 

```{r}
actors_test<-mv_test%>%
  group_by(actor1)%>%
  count()%>%
  arrange(-n)%>%
  ungroup()%>%
  slice(1:50)%>%
  select(actor1)%>%
  as_vector()
  

actors_train<-mv_train%>%
  group_by(actor1)%>%
  count()%>%
  arrange(-n)%>%
  ungroup()%>%
  slice(1:50)%>%
  select(actor1)%>%
  as_vector()

actor_list<-actors_test[actors_test%in%actors_train]

for (actor in actor_list){  

actor_name<-paste0(str_remove_all(actor, " "),"_ac")
  
mv_train<-mv_train%>%
  mutate(temp =ifelse(str_detect(actor1,actor),1,0))
  
mv_train[actor_name]<-mv_train$temp

mv_train<-mv_train%>%select(-temp)
}
  


for (actor in actor_list){  

actor_name<-paste0(str_remove_all(actor, " "),"_ac")
  
mv_test<-mv_test%>%
  mutate(temp =ifelse(str_detect(actor1,actor),1,0))
  
mv_test[actor_name]<-mv_test$temp

mv_test<-mv_test%>%select(-temp)
}

```

## Run time

Run time was not originally in numeric format, but with a bit of work we can get it there. 

```{r}
mv_train<-mv_train%>%
  mutate(runtime=parse_number(runtime))

mv_test<-mv_test%>%
  mutate(runtime=parse_number(runtime))

```


## Timing of Release

Movies released at different times of year might do better. Below I include the month of the release as a nominal variable to include in the model.

```{r}
mv_train<-mv_train%>%
  mutate(rel_date=lubridate::ymd(released))%>%
  mutate(rel_month=as.character(lubridate::month(rel_date)))


mv_test<-mv_test%>%
  mutate(rel_date=lubridate::ymd(released))%>%
  mutate(rel_month=as.character(lubridate::month(rel_date)))

```


```{r}
mv_train<-mv_train%>%
  mutate(year=as.character(year))

mv_test<-mv_test%>%
  mutate(year=as.character(year))
```


## Select desired features

I'm going to now subset to just the variables I want in the analysis. 

```{r}
mv_train_select<-mv_train%>%
  select(imdbid,
         log_box_office,
         imdbrating,
         imdbvotes,
         runtime,
         metascore,
         matches("_wd"),
         matches("_ac"),
          language1,
          production1,
         year,
          rated,
          oscar_winner,
         country1,
         writer1,
          genre1,
          rel_month
          )%>%
  drop_na()%>%
  mutate(year=as.character(year))%>%
  mutate(rel_month=as.character(rel_month))


mv_test_select<-mv_test%>%
  select(imdbid,
         imdbrating,
         imdbvotes,
         runtime,
         metascore,
         matches("_wd"),
         matches("_ac"),
         language1,
         production1,
         year,
         rated,
         oscar_winner,
         genre1,
         country1,
         writer1,
         rel_month
         )%>%
  drop_na()%>%
  mutate(year=as.character(year))%>%
  mutate(rel_month=as.character(rel_month))

```


## Set Recipe

```{r}
mv_formula<-as.formula("log_box_office~.")

mv_rec<-recipe(mv_formula,mv_train_select)%>%
  update_role(imdbid,new_role = "id variable")%>%
  update_role(log_box_office,new_role="outcome")%>%
  step_naomit(all_predictors())%>%
  step_other(all_nominal_predictors() )%>%
  step_dummy(all_nominal_predictors())%>%
  step_log(imdbvotes)%>%
  step_zv(all_predictors())%>%
  step_normalize(all_predictors())
```

## Check Prepped Training Data

```{r}
prep_data<-prep(mv_rec)%>%bake(mv_train)
```

## Structure Data for Resampling

```{r}
mv_train_select_rs<-mc_cv(mv_train_select)
```

## Set Model
```{r}
lasso_fit<-linear_reg(penalty = tune(),
                     mixture=1)%>%
  set_engine("glmnet")%>%
  set_mode("regression")
```


## Create Workflow
```{r}
mv_wf<-workflow()%>%
  add_model(lasso_fit)%>%
  add_recipe(mv_rec)
```


## Create the Tuning Grid
```{r}
lasso_grid<-grid_regular(parameters(lasso_fit) ,levels=10)
```



## Tune the Model
```{r}
movie_lasso_tune_fit <- 
  mv_wf %>%
    tune_grid(mv_train_select_rs,grid=lasso_grid)
```

## Check Model Fit 
```{r}
movie_lasso_tune_fit%>%
  collect_metrics()%>%
  filter(.metric=="rmse")%>%
  arrange(mean)
```


## Select Best Fit
```{r}
movie_final<-
  finalize_workflow(mv_wf,
                    select_best(movie_lasso_tune_fit,
                                metric="rmse"))%>%
  fit(mv_train_select)
```


## Examine Coefficients in Model of Best Fit
```{r}
movie_final%>%
  tidy()%>%
  filter(estimate>.01)%>%
  arrange(-estimate)%>%
  View()
```

## Predict in Testing Dataset
```{r}
mv_test_select<-
  movie_final%>%predict(mv_test_select)%>%
  bind_cols(mv_test_select)%>%
  rename(pred=.pred)%>%
  select(imdbid,pred)
```

```{r}
write_csv(mv_test_select,"group_prof.csv")
```


