---
title: "Topic 6. Conditional Variation"
author: "Josh Clinton"
date: "9/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Today's Agenda

- Becoming political pundits: summarizing the 2020 Popular Vote Over Time

- Conditional data: when a variable varies with respect to some other variable

- Visualizing conditional data

- "Smoothing" data 

- (Intro) Looping

#  What is our question?

How did the support for Biden and Trump vary across the course of the 2020 Election?

- What should we measure?

- How do we summarize, visualize, and communicate?

- We want to evaluate the properties of presidential polling as Election Day 2020 approached.

- Necessary for prediction -- we want most recent data to account for last-minute shift.

- Necessary for identifying when changes occurred (and why?)

# Dates in R

- Dates are a special format in R (character with quasi-numeric properties)

```{r}
load(file="data/Pres2020.PV.Rdata")
election.day <- as.Date("11/3/2020", "%m/%d/%Y")   
election.day16 <- as.Date("11/8/2016", "%m/%d/%Y")   
```

- Difference in "dates" versus difference in integers?

```{r}
election.day - election.day16
as.numeric(election.day - election.day16)
```

## Conditional Relationships

- How does the value of the outcome of interest vary _depending_ on the value of another variable of interest?

- Outcome of interest (dependent variable, Y)

- Other variables possibly related to the outcome (independent variable, X)

Y: Number of Polls being reported on
X: Proximity to Election Day

So, for every day, how many polls were reported by the media?

## Let's Wrangle...

```{r}
Pres2020.PV <- Pres2020.PV %>%
                mutate(EndDate = as.Date(Pres2020.PV$EndDate, "%m/%d/%Y"), 
                      StartDate = as.Date(Pres2020.PV$StartDate, "%m/%d/%Y"),
                      DaysToED = as.numeric(election.day - EndDate),
                      margin = Biden - Trump)
```

## Final Barplot

```{r}
p <- ggplot(data = Pres2020.PV, aes(x = DaysToED)) +
  labs(title = "Number of 2020 National Polls Over Time") + 
  labs(x = "Number of Days until Election Day") + 
  labs(y = "Count") + 
  geom_bar(color="PURPLE")  + 
  scale_x_reverse(breaks=seq(0,230,by=10))

p
```



## Black and White Histogram

- Is this better?

- How does a bin in a histogram compare to a bar in a barplot?

```{r}
p <- ggplot(data = Pres2020.PV, aes(x = DaysToED)) +
  labs(title = "Number of 2020 National Polls Over Time") + 
  labs(x = "Number of Days until Election Day") + 
  labs(y = "Count") + 
  geom_histogram(color="PURPLE",bins = 30) +
  scale_x_reverse() + 
  theme_bw()

p
```


#   Bivariate/Multivariate relationships

- Most of what we do is a relationship between (at least) 2 variables.

- Here we are interested in how the margin varies as Election Day approaches: `margin` by `DaysToED`.

- Want to plot X (variable that "explains") vs. Y (variable being "explained"):

# Scatterplot: Relationship between Continuous Variables

```{r}
margin.plot <- ggplot(Pres2020.PV, 
                      aes(x = EndDate, y = margin)) + 
  labs(title="Margin in 2020 Nat. Popular Vote Polls Over Time") + 
  labs(y = "Margin: Biden - Trump") + 
  labs(x = "Poll Ending Date")+ geom_point(color = "PURPLE")

margin.plot
```

## Two things too change

1. Axis looks weird - lots of interpolation required by the consumer.



- `scale_x_date` vs. `scale_y_continuous` - when use?

```{r}
margin.plot  + 
    geom_point(color = "PURPLE")  + 
    scale_y_continuous(breaks=seq(-10,20,by=5)) +
    scale_x_date(date_breaks = "2 week", date_labels = "%b %d") 

margin.plot
```

2. Data looks "chunky"?  How many data points are at each point?


- Using `jitter` - what does it do?  Why not the best?

```{r}
margin.plot  + 
    geom_point(color = "PURPLE", position="jitter") + 
    scale_y_continuous(breaks=seq(-10,20,by=5)) +
    scale_x_date(date_breaks = "2 week", date_labels = "%b %d") 

margin.plot
```

- What does `alpha` do?

```{r}
margin.plot  + 
    geom_point(color = "PURPLE", alpha = .4) + 
    scale_y_continuous(breaks=seq(-10,20,by=5)) +
    scale_x_date(date_breaks = "2 week", date_labels = "%b %d") 

margin.plot
```


```{r}
margin.plot  + 
  geom_point(color = "PURPLE", alpha = .4) + 
    scale_y_continuous(breaks=seq(-10,20,by=5)) +
    scale_x_date(date_breaks = "2 week", date_labels = "%b %d")  + geom_line()
```

## "Smoother" lines?

- `geom_smooth` adds in a weighted ("smoothed") average

- Better, but do you understand a loess regression?

```{r}
margin.plot  + 
  geom_point(color = "PURPLE", alpha = .4) + 
    scale_y_continuous(breaks=seq(-10,20,by=5)) +
    scale_x_date(date_breaks = "2 week", date_labels = "%b %d")  + 
  geom_smooth(color = "BLACK", se=F) 
```

RECAP:

- `geom_point` adds a point at the `(x,y)` point defined in `ggplot` (unless defined in `geom_point`)
- `geom_line` adds a line connecting the `(x,y)` points


# Plotting Multiple Variables Over Time (Time-Series)

- Can we plot support for Biden and support for Trump separately over time (on the same plot)?

OR

- Comparing the change in `margin` over time for multiple election years?

- Comparing the support for candidates (`Biden` and `Trump`) in multiple states?

- Comparing the support for candidates according to different types of polls?

- Comparing the support for presidential candidates relative to senatorial and gubernatorial candidates in the same state?

- Comparing the deaths/cases per capita over time (and also by county/state)?

- Comparing the performance of an NBA team/player in several dimensions over time?

## Plotting two variables from same tibble on same ggplot

```{r}
BidenTrumpplot <- ggplot(Pres2020.PV)  +
  labs(title="% Biden and Trump in 2020 National Popular Vote Polls Over Time") + 
  labs(y = "Pct. Support") + 
  labs(x = "Poll Ending Date") + 
  geom_point(aes(x = EndDate, y = Trump), 
             color = "red", alpha=.4)  +
  scale_x_date(date_breaks = "2 week", date_labels = "%b %d") + 
  geom_point(aes(x = EndDate, y = Biden), 
             color = "blue", alpha=.4)

BidenTrumpplot
```

- Which axis do you prefer of the following? Why?

```{r}
BidenTrumpplot + ylim(0,100)
```

```{r}
BidenTrumpplot + scale_y_continuous(breaks=seq(30,70,by=5))
```

## Putting it all together

```{r}
BTNational <- ggplot(Pres2020.PV) + 
  geom_point(aes(x = EndDate, y = Trump), 
             color = "red", alpha = .4) + 
  geom_point(aes(x = EndDate, y = Biden), 
             color = "blue", , alpha = .4)  + 
  geom_smooth(aes(x = EndDate, y = Trump), 
              color = "red",se=F) + 
  geom_smooth(aes(x = EndDate, y = Biden), 
              color = "blue",se=F) +   
  labs(title="% Biden and Trump in 2020 Nat. Popular Vote Polls Over Time") + 
  labs(y = "Pct. Support") + 
  labs(x = "Poll Ending Date") + 
  scale_x_date(date_breaks = "2 week", date_labels = "%b %d") + 
  scale_y_continuous(breaks=seq(30,70,by=5))

BTNational
```


## "Smoothing"

- How can we understand the role of "smoothing"?

- Very important because that is the trend we focus on!

- Cannot rely on a default we do not understand.

- We want to summarize the trend by looking at the polling average for all polls conducted in a certain time period (bandwidth).

- Smoothing helps ensure that an outlier does not distort our visualization (and our interpretation of central tendency of the data)

- Smoothing is taking a mean _conditional_ on some other set of values -- here time!


## Conditioning Variable: Time

Start by defining a variable `all_dates` -- all possible dates of interest (not just those with data!)

```{r}
all_dates <- seq(min(Pres2020.PV$EndDate), election.day, 
                 by = "days")
```

# GOAL

- For each possible date, what is the average support for Biden and Trump among the polls taken during the `X` days prior?

- Requires define a moving "bandwidth" of dates and calculating average support among polls conducted in that bandwidth.

- To do this we are going to "loop" over dates. (ASIDE: Very inefficient; vectorize your computations if possible!)

# Looping

GOAL: Calculate the average support for Trump in the last 3 days of the election.

- I could `filter`, `group_by`, and `summarize` but I could also:

```{r,eval=FALSE}
mean(Pres2020.PV$Trump[Pres2020.PV$EndDate == "2020-10-31"])
mean(Pres2020.PV$Trump[Pres2020.PV$EndDate == "2020-11-01"])
mean(Pres2020.PV$Trump[Pres2020.PV$EndDate == "2020-11-02"])
```

- Could do this for every day, but: very inefficient & prone to error (copy and paste bad!)!

# Looping
```{r}
dates <- c("2020-10-31","2020-11-01","2020-11-02")
```

Format of a loop is:

```{r, eval=FALSE}
for(i in dates){
  CODE TO REPEAT HERE
}
```

# What does this do?

```{r}
for(i in dates){
  print(i)
}
```


# What does this do?

```{r}
for(i in dates){
  print(i)
  mean(Pres2020.PV$Trump[Pres2020.PV$EndDate == i])
}
```

- printing `i` Useful for debugging!

# What does this do?

```{r}
PollAvg <- NULL

for(i in dates){
  print(i)
  PollAvg[i] <- mean(Pres2020.PV$Trump[Pres2020.PV$EndDate == i])
}

PollAvg
```

# Preliminaries for the Loop

```{r}
Bandwidth <- 2
PV_avg <- vector(length(all_dates), mode = "list") 

for (i in seq_along(all_dates)) {
  date <- all_dates[i]

    PV_avg[[i]] <- Pres2020.PV %>%
    filter(as.integer(EndDate - date) <= 0 &
           as.integer(EndDate - date) > - Bandwidth) %>%
    summarize(Biden = mean(Biden),
              Trump = mean(Trump)) %>%
      mutate(date = date)
}

pop_vote_avg <- bind_rows(PV_avg)
```


- QUESTION: Why didn't I do what we did when analyzing the NBA data and use `sample_n`?  What is different here versus there?

## Plot with 2 day smoother

- Why gaps?

```{r}
PlotTS <- ggplot(pop_vote_avg) + 
  geom_line(aes(x = date, y = Trump), color = "red") + 
  geom_line(aes(x = date, y = Biden), color = "blue")  + 
  labs(title="2-Day Avg. Support for Biden and Trump in 2020 National Popular Vote Polls Over Time") + 
  labs(y = "Avg. Support") + 
  labs(x = "Poll Window Ending Date (2 Days)") + 
  scale_x_date(date_breaks = "2 week", date_labels = "%b %d") + 
  scale_y_continuous(breaks=seq(30,70,by=5))

PlotTS
```


# Weeklong bandwidth?

```{r}
Bandwidth <- 7
PV_avg <- vector(length(all_dates), mode = "list") # holding variable

for (i in seq_along(all_dates)) {
  date <- all_dates[i]
  
  PV_avg[[i]] <- Pres2020.PV %>%
    filter(as.integer(EndDate - date) <= 0,
           as.integer(EndDate - date) > - Bandwidth) %>%
    summarize(Biden = mean(Biden),
              Trump = mean(Trump)) %>%
    mutate(date = date)
}

pop_vote_avg7 <- bind_rows(PV_avg)
```

Now plot!

## Plotting four variables from two tibbles on same ggplot!


```{r}
# Now overlay on points!
PopVotePlot <- ggplot() + 
  geom_point(data=Pres2020.PV,aes(x = EndDate, y = Trump), 
             color = "pink", alpha=.4) + 
  geom_point(data=Pres2020.PV,aes(x = EndDate, y = Biden), 
             color = "light blue", alpha=.4) + 
  geom_line(data=pop_vote_avg7, aes(x = date, y = Trump), 
            color = "red") + 
  geom_line(data=pop_vote_avg7, aes(x = date, y = Biden), 
            color = "blue") + 
  labs(title="Support for Biden and Trump in 2020 National Popular Vote Polls Over Time") + 
  labs(y = "Pct. Support") + 
  labs(x = "Poll Ending Date") + 
  scale_x_date(date_breaks = "2 week", date_labels = "%b %d") + 
  scale_y_continuous(breaks=seq(30,70,by=5))

PopVotePlot
```

# Going Forward: On your own?

- What is the right bandwidth?  How much change is "real"?

- Every poll counted equally (`SampleSize`)?

- Every type of poll counted equally? (`filter` by different types of polls?)

- Is this really what we care about? National Popular Vote?

- How deal with uncertainty?

- Other data? (Past?  Current? Entirely different?)

