---
title: "Topic8 Mapping Electoral College Predictions"
author: "Josh Clinton"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


#   Today

- Maps in `ggplot` (one way).
- Pulling it together.  State level predictions and Electoral College.

Our data this time is all state-level presidential polls conducted in 2020.

```{r}
load(file = "data/StatePresidentialVote2020.Rdata")
glimpse(dat)
```

Produce a summary by state by averaging across all polls that are done in a state.  

```{r}
State.mean <- dat %>%
  group_by(region) %>%
      summarize(BidenPct = mean(Biden, na.rm = TRUE),
                TrumpPct = mean(Trump, na.rm = TRUE),
                NumPolls = n())


State.mean
```


# Maps in R

There are several packages that let you do maps in R.  Let us use one called `maps`.

```{r}
#install.packages("maps")   # Install this if you don't already have it!
library(maps) 
```

- This package contains the data needed to produce a map that we extract using the funciton `map_data`.
- The dimensions of `states48` are large because this is an object that contains the points of latitude and longitude needed to draw the map.

```{r}
states48 <- map_data("state")
dim(states48)
states48[1,]
```

# Creating the canvas

- The map is drawn using `geom_polygon` which uses the `long` and `lat` to draw the points in `ggplot`. 
- Note that the default is pretty ugly (and distracting)

```{r}
ggplot() + 
  geom_polygon(data=states48, aes(x=long, y=lat, group=group),
                color="black", 
                fill="lightblue" )
```

Let's clean in up and save it as an object to be used later.  (Note that we can create a custom theme to apply to a `ggplot`)

```{r}
plotmap <- ggplot() + 
  geom_polygon(data=states48, aes(x=long, y=lat, group=group),
                color="black", 
                fill="lightblue" ) + 
  labs(y = "") + 
  labs(x = "") + scale_x_continuous(breaks=NULL, labels=NULL) + 
  scale_y_continuous(breaks=NULL, labels=NULL) + 
    theme(
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    )

plotmap
```

# Get Some Data!
```{r}
states48.join <- inner_join(states48,State.mean)
states48.join[1,]
```

- `inner_join()` merge dataframes if common in both (must have same name)

- `outer_join()` merge dataframes if in either

Produce a map that indicates predicted winners depending on the average polling done in each state using different polygons.

```{r}
plotmap + 
  geom_polygon(data=states48.join, 
               aes(x=long, y=lat, group=group)) + 
  geom_polygon(data=filter(states48.join,BidenPct > TrumpPct), 
               aes(x=long, y=lat, group=group),fill="Blue") +
  geom_polygon(data=filter(states48.join,BidenPct < TrumpPct), 
               aes(x=long, y=lat, group=group),fill="Red") 
```

But we can do the same thing using `fill` -- here defined to be which candidate is larger.

```{r}
statewinner <- plotmap + 
  geom_polygon(data=states48.join, 
               aes(x=long, y=lat, group=group,
                   fill=BidenPct > TrumpPct),
               color="white", size = .1) 

statewinner
```

- We can also add more nuanced information by denoting the location of states where the margin is less than 2 -- i.e., very close.
- Because we are adding on top of the canvass we are just "repainting" the impacted states.

```{r}
statewinner +
  geom_polygon(data=filter(states48.join,
               BidenPct > TrumpPct & BidenPct - TrumpPct < 2), 
               aes(x=long, y=lat, group=group),
               fill="light blue") + 
  geom_polygon(data=filter(states48.join,
               BidenPct < TrumpPct & BidenPct - TrumpPct > 2), 
               aes(x=long, y=lat, group=group),
               fill="light red") 
```

## IN-CLASS: Can you plot the location of "Blow out" states?  How do you define that?



# Using Fill as a variable

- Instead of using different colors to define categories of difference we can also use the `fill` to try to convey information about the variation of continuous differences -- e.g., the level of support.

```{r}
plot1 <- plotmap + 
  geom_polygon(data=states48.join, 
               aes(x=long, y=lat, group=group,
                   fill=BidenPct/100),
               color="WHITE", size = .1)

plot1
```

And we can also define the scale in terms of the color associated with the min and max values.
```{r}
plot1 + scale_fill_continuous(name="Biden Pct", 
                      low = "Red", high = "Blue")
```

- But how much better is this than a histogram?  Does the geography really matter here?  

- Why a histogram and not a barplot?  What would a barplot look like?

```{r, echo=FALSE, warning=FALSE}
State.mean %>% ggplot() + 
  scale_x_continuous(breaks= seq(30,70,by=5)) +
  labs(x="Biden Polling Average in State",
       y="Number of States",
       title="2020 Biden Overall Polling Average Across States") +
    geom_histogram(mapping = aes(BidenPct), fill = "blue", binwidth=2.5)
```
