---
title: "Topic 8. Electoral College Predictions"
author: "YOUR NAME"
output: html_document
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 5)
library(tidyverse)
```

Load the data of all state-level polls for the 2020 presidential election.

```{r}
load(file="data/Pres2020.StatePolls.Rdata")
glimpse(Pres2020.StatePolls)
```


Variables of potential interest include:

- _Biden_ : Percentage of respondents supporting Biden, the Democrat, in poll (0-100)
- _Trump_ : Percentage of respondents supporting Trump, the Republican, in poll (0-100)
- _BidenCertVote_ : Percentage of vote Biden, the Democrat, actually received in the election (0-100)
- _TrumpCertVote_ : Percentage of vote Trumo, the Democrat, actually received in the election (0-100)
- _Winner_ : whether Biden won ("Dem") or Trump won ("Rep")
- _poll.predicted_ : Indicator for whether the poll correctly predicted who won (1) or not (0)
- _State_ & _StateName_ : the state where the poll was conducted
- _EV_ : the number of Electoral College Votes the state is worth for the winning candidate

# Overall Task:  

- How do we use this data to calculate the probability that Biden will win the presidency of the United States by winning the Electoral College?

### Task 1: How should we translate a poll result into a predicted probability?

Suppose that I give you 10 polls from a state.  How can you use them to create a probability?  Discuss!  (I can think of 3 ways.)


### Task 2: Start "simple" -- calculate the probability that Biden wins PA

- Create a tibble with just polls from PA.

```{r}

```

- Now compute the three probabilities.  What functions do we need?

```{r}

```

- What do you think about the differences?

### Task 3: Given those probabilities, how do we change the code to compute the expected number of Electoral College Votes `EV` for Biden?

- Keep the code from above and copy and paste so you can understand how each step changes what we are doing.

```{r}


```
### Task 3B: How get expected Electoral College votes for Trump?


- NOTE: There are at least 2 ways to do this because this is a 2 candidate race


### Task 4:  Now generalize to every state by applying this code to each set of state polls.

- What do we need to do this calculation for every state in our tibble?
- First, compute probability of winning a state. (How?)
- Second, compute expected Electoral College Votes. (How?)

```{r}

```

### Task 5:  Now compute total expected vote by adding to that code

- NOTE: Actually 306 - 232
- What do we need to do to the tibble we created in Task 4 to get the overall number of Electoral College Votes?

```{r}
```


- You could also do this for just polls conducted in the last X days.  How?


# So far we have taken polls as fixed. What if we allow randomnesss via resampling?

###  Start "simple": probability that Biden wins PA

Write a loop that, for each iteration....

- Draws a sample of the same number of polls as were done in PA using `sample_n`
- Calculates the probability of a win using measures just discussed

But before you start, how does this differ from what we did previously with the PA data?  What do we need to add to the code?

```{r}
ProbBidenWin.PA <- NULL
NSamples <- 1000

for(i in 1:NSamples){

  }
```

### Compute probability that Biden wins and the range containing 95% of the expected means (if we were to redo everything exactly as is).

```{r}

```

### Can you do this for Wisconsin?

- Take it from start to finish for a different state

```{r}

```

## Now do this for all the states!

- Let us draw a new sample of 1545 state-level polls and go...

```{r}
ElectoralCollegeVotes <- NULL
B <- 1000

for(i in 1:B){

  }
```

### So what is the probability Biden gets at least 270 Electoral College Votes?

```{r}
 
```
### Now plot?

```{r}

```

### Thinking

- There are lots of things we could change here!