---
title: "Univariate Visualization Using `ggplot`"
author: "YOUR NAME HERE"
date: 9/21/21
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#     Principles of Data Visualization: Do's and Don'ts

INSERT NOTES/THOUGHTS HERE

#     US COVID Map

A popular representation of the pandemic is: https://coronavirus.jhu.edu/us-map

I think this is a good/bad visualization because....

#   Application National Popular Vote in 2020

We are going to analyze data that...

First load the data.  Note that the `Rdata` object is in a folder called "data".  If you want to put it in the same folder just delete that part.

```{r}
load(file="data/Pres2020.PV.Rdata")
glimpse(Pres2020.PV)
```

#   Define some variables we need

Create a variable called `margin` for Biden - Trump using `mutate`

```{r}
Pres2020.PV <- Pres2020.PV %>%
  mutate()
```

#   Barplots

We use a barplot for....

The code to create a barplot using `ggplot` is....

```{r}

```

# Everything is an object!

Adding variable labels and a vertical line summarizing the mean:

```{r}

```

# Saving the plot

You can save the plot manually using RStudio and the "Plots" window, but better is to program it directly.

Here is how we save a plot:

```{r}

```


#   Histograms

We use histograms for....

The default histogram code is...

```{r}
```

#   Tweaking Histograms

We can change the number of "bins" being plotted as follows... 

```{r}
```

#     Advanced: Density Plots

A density plot is....

It can be useful for...

We can convert a histogram plot to a density plot using....

```{r}
```


#   On your own

Try different things!  
- Change the color?  
- Change the number of bins?
- Save the graph as a pdf?
- Plot some different variables: `Trump`, `Biden`, `Mode`, `SampleSize`



