---
title: "Assignment 2"
author: "YOUR NAME HERE"
date: "`r Sys.Date()`"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
```


Submit a knitted HTML and the orignal *.Rmd file named using course naming conventions: `LASTNAME_FIRSTINITIAL_Assignment2`.

Grading: 15 points total. (19 with BONUS.) 1 point for correct submission and markdown format.  1 point each for correct code chunks for parts 1-9.  2 points for correct code for question 10. 1 point total for correctly interpreting results for questions 4 and 9.  1 point for correct explanation of the tibble produced in question 8.  1 point for the interpretation of the question 10.


1.  Load in the data `ExitPolls.Rds`

```{r}
ExitPolls <- readRDS( file = "data/ExitPolls.Rds")
```

2. Create a new tibble named `ExitPolls_small` that contains only the variables: `female`, `AgeUnder30`, `Age6074`, `state`, and `vote.choice` and which removes any respondent who fails to vote for Biden or Trump.

```{r}
ExitPolls_small <- ExitPolls %>%
  select(female,AgeUnder30,Age6074,state,vote.choice) %>%
  filter(vote.choice == "Biden" | vote.choice == "Trump")
```

3. Take a glimpse of the resulting data

```{r}
glimpse(ExitPolls_small)
```

4.  Using the `count` function, look at the distribution of values for `female`.  What do you observe?
```{r}
ExitPolls_small %>%
  count(female)
```

5. Write code that drops missing values for `female` in `ExitPolls_small` and saves the result.
```{r}
ExitPolls_small <- ExitPolls_small %>%
  drop_na(female)
```    

6. Create a new variable named `BidenVoter` in the tibble `ExitPolls_small` that takes on the value of 1 for Biden voters and 0 for Trump voters.  To do so use `mutate` and `ifelse`.

```{r}
ExitPolls_small <- ExitPolls_small %>%
  mutate(BidenVoter = ifelse(vote.choice=="Biden",1,0))
```

7. Using the `summarize` function, write code that reports the proportion of respondents under the age of 30 who report voting Biden. 

```{r}
ExitPolls_small %>%
  filter(female ==1 & AgeUnder30==1) %>%
  summarize(PctBiden = mean(BidenVoter))
```

8. But suppose that we want to make comparisons across both age and sex. Use the variables `female` and `AgeUnder30` to create four tibbles -- one for each combination of age and sex -- and compute the proportion of `BidenVoter` for each.  Compare the differences by sex within age.  Then compare difference by age within sex.  How do you interpret what you observe?

```{r}
# Create the Tibbles
FemaleUnder30<- ExitPolls_small %>%
  filter(female==1 & AgeUnder30 == 1) %>%
  summarize(PctBiden = mean(BidenVoter))

MaleUnder30 <- ExitPolls_small %>%
  filter(female==0 & AgeUnder30 == 1) %>%
  summarize(PctBiden = mean(BidenVoter))

FemaleOver30 <- ExitPolls_small %>%
  filter(female==1 & AgeUnder30 == 0) %>%
  summarize(PctBiden = mean(BidenVoter))

MaleOver30 <- ExitPolls_small %>%
  filter(female==0 & AgeUnder30 == 0) %>%
  summarize(PctBiden = mean(BidenVoter))
```

Now with this tibbles compare the percent who support Biden by sex for younger voters in theses states and older voters.  We observe roughly similar differences between female and male voters with females being more likely to support Biden than Trump, although the gap is slightly bigger among younger voters.

```{r}
FemaleUnder30 - MaleUnder30
FemaleOver30 - MaleOver30
```

Now compare older and younger females and males to look at the variation by age. Here we see much bigger differences than we saw before. This suggests that age differences are larger than sex differences in the data we have.  In fact, the differences by age are nearly twice as large as the differences by sex.  This suggests that perhaps the generational divide is a more important cleavage in American politics than the ``gender divide."  It also has strong consequences given that younger voters do not vote as much as older voters and this matters _a lot_.

```{r}
FemaleUnder30 - FemaleOver30 
MaleUnder30 - MaleOver30
```

9.  But maybe this varies by state? To start, how many respondents under the age of 30 do we have in Florida (FL), North Carolina (NC), Pennsylvania (PA), and Texas (TX)?  

```{r}
ExitPolls_small %>%
  filter(AgeUnder30==1) %>%
  count(state) 
```

10. The data contains respondents from four states: FL, NC, PA, and TX. Compare the difference in support for Biden and Trump for male and female voters in the states of North Carolina (NC) and Pennsylvania (PA).  What do you observe?  Which group is most and least supportive of Biden?  Given the results of question 9 how comfortable are you with the results?

For North Carolina, we see a very small difference of around 3 points with 65\% of males and 68\% of females under the age of 30 supporting President Biden.

```{r}
ExitPolls_small %>%
  filter(AgeUnder30==1 & state=="NC" & female==0) %>% 
    summarize(PctBiden = mean(BidenVoter))       

ExitPolls_small %>%
  filter(AgeUnder30==1 & state=="NC" & female==1) %>% 
    summarize(PctBiden = mean(BidenVoter))       
```

For Pennsylvania we see a much larger gap.  63\% of males supported President Biden but nearly 82\% of females under the age of 30 reported supporting President Biden.  This is a massive gap of 20 percentage points.  However, given that there are only 166 respondents under the age of 30 in PA, but there are 455 in NC, we do worry that our data may be unrepresentative as it seems odd that one survey is so much smaller than the other one.  The small sample size and the fact that the gap is so large compare to NC (and also the overall gap we found in question 8) makes me wary of reading too much into the results.
         
```{r}
ExitPolls_small %>%
  filter(AgeUnder30==1 & state=="PA" & female==0) %>% 
    summarize(PctBiden = mean(BidenVoter))       

ExitPolls_small %>%
  filter(AgeUnder30==1 & state=="PA" & female==1) %>% 
    summarize(PctBiden = mean(BidenVoter))       
```

BONUS (+4 POINTS POSSIBLE). Using the data we used in class `MI_final_small` compare the importance of racism between Democrats and Republicans (+1) to the importance between White and Black voters (+2).  What do you observe and why do you think that is (+1)?

```{r}
# INSERT CODE HERE
```

