---
title: "Prediction in Testing Data"
author: "Will Doyle"
date: "11/1/2021"
output: html_document
---

A quick tip: you need to make sure your training and testing data are EXACTLY the same before you try and generate predictions in the testing data. This means that any data wrangling you do on the training data must be done on the testing data. There are a couple of ways to do this, I detail one below.

```{r}
library(tidyverse)
library(tidymodels)
```

I want to make sure that the numeric variables are included as numeric in the training and testing data. The numeric variables include "N/A" for missing data, which I don't want. So I have two steps: I'm going to convert N/A to NA then convert `boxoffice` and `metascore` to numeric. I have to do the exact same thing to the testing data. The only difference is that there's no `boxoffice` variable in the testing data. 

```{r}
## Wrangling: _exactly_ the same for both

train<-read_rds("movie_train.Rds")%>%
  mutate(boxoffice_a=ifelse(boxoffice_a=="N/A",NA,boxoffice_a))%>%
  mutate(metascore=ifelse(metascore=="N/A",NA,metascore))%>%
  mutate(across(.cols  =c(boxoffice_a,metascore), as.numeric))%>%
  filter(!is.na(boxoffice_a))%>%
  mutate(log_box_office=log(boxoffice_a))


test<-read_rds("movie_test.Rds")%>%
  mutate(metascore=ifelse(metascore=="N/A",NA,metascore))%>%
  mutate(across(.cols=c(metascore),as.numeric))
```


From there, I can run the normal steps, first I specify a model:

```{r}
lm_fit<-linear_reg()%>%
  set_engine("lm")%>%
  set_mode("regression")
```

Then a formula: 

```{r}
mv_formula<-as.formula("log_box_office~metascore+language1")

mv_recipe<-recipe(mv_formula,train)%>%
  step_other(language1)%>%
  step_dummy(language1)

```


I add both of those to a workflow:
```{r}
mv_wf<-workflow()%>%
  add_model(lm_fit)%>%
  add_recipe(mv_recipe)
```

Then I fit my model to the data:
```{r}
mv_wf<-mv_wf%>%
  fit(train)
```

Take a look at the results
```{r}
mv_wf%>%tidy()
```

Predict in the testing dataset
```{r}
test<-mv_wf%>%
  predict(test)%>%
  bind_cols(test)%>%
  select(imdbid,.pred)
```

And create a .csv file for submission
```{r}
write_csv(test,"submit.csv")
```

