---
title: "Most of the Functions We Have Talked About"
date: "10/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Here is a document to get you started.  You should add to this as the course goes along. `ggplot` functions are not listed.  Note that there are often other toggles that can be used in each function.  When in doubt you can either use `help(FUNCTION)` where `FUNCTION` is the name of what you are trying to use or else use the internet.  There are literally hundreds of guides, e.g., https://www.rstudio.com/resources/cheatsheets/

# `load`

- Syntax: `load(FILE)`
- This will load an R-formatted dataframe/tibble into the Global Environment.  Note that the name of the object being loaded is the object that it was saved as, not the filename being loaded.
```{r}
load(file="data/Pres2020.PV.Rdata")
```

# `arrange`
- Syntax `arrange(TIBBLE,VARIABLE)`
- Reorders the rows of the `TIBBLE` according to the values of `VARIABLE` from smallest to largest.  Does not change any of the values.
- NOTE: `-VARIABLE` arranges from largest to smallest
- NOTE: that if we run `arrange` in a "pipe" then we do not need to explicitly name `TIBBLE` as the pipe will identify its' relationship.

```{r}
arrange(Pres2020.PV,SampleSize)

Pres2020.PV %>%
    arrange(-SampleSize)
```

# `count`
- Syntax: `count(TIBBLE,VARIABLE1)`
- Produces a count of the number of elements in each value of the named variable.  
- NOTE: that if we run `count` in a "pipe" then we do not need to explicitly name `TIBBLE` as the pipe will identify its' relationship.


```{r}
count(Pres2020.PV,Mode)

Pres2020.PV %>%
    count(Mode) %>%
    arrange(-n) # Arrange in decreasing order
```

# `filter`
- Syntax `filter(TIBBLE,CONDITION)`
- Selects the rows of the TIBBLE according to whether `CONDITION` is satisfied.
- NOTE: `CONDITION1 & CONDITION2` means both conditions must be satisfied (AND) ;  `CONDITION1 | CONDITION2` means either conditions must be satisfied (OR) 
- NOTE: that if we run `filter` in a "pipe" then we do not need to explicitly name `TIBBLE` as the pipe will identify its' relationship.

```{r}
filter(Pres2020.PV, Mode=="Online" & !is.na(MoE))

OnlinePolls <- Pres2020.PV %>%
                filter(Mode == "Online" & !is.na(MoE))
```

# `group_by`
- Syntax `group_by(VARIABLE1,VARIABLE2)`
- Defines the grouping on which a set of subsequent computations will be done on.
- Groups are defined in the TIBBLE sorting by `VARIABLE1`, then by `VARIABLE2`, etc.
- NOTE: To stop calculations being done done within the grouping we defined we need to `ungroup`  

```{r}
Pres2020.PV %>%
    group_by(Mode,SampleSize) %>%
    count() %>%
    ungroup()
```


# `ifelse`
- Syntax: `ifelse(CONDITION,OUTPUT1,OUTPUT2)`
- Creates a variable that takes on the value `OUTPUT1` if `CONDITION` is true, and `OUTPUT2` if it is not.


```{r}
DATA.VECTOR <- seq(1,8)
ifelse(DATA.VECTOR > 5, "Greater than 5", "Less than or equal to 5")
```

#  `length`
- Syntax: `length(DATA.VECTOR)` 
- `length` returns the number of observations in the vector `DATA.VECTOR`.  It is a number.

```{r}
length(DATA.VECTOR)
```


# `mean`

- Syntax: `mean(DATA.VECTOR, na.rm=FALSE)` 
-  Calculates the sample average of the vector `DATA.VECTOR`. It is a number.  
- `na.rm=FALSE` means that missing values will cause a failure to compute the means.
- `na.rm=TRUE` calculates the mean after removing missing data.

```{r}
DATA.VECTOR <- c(1,2,3,NA,3)
mean(DATA.VECTOR)
mean(DATA.VECTOR, na.rm = TRUE)
```

# `sample`

- Syntax: `sample(x = DATA.VECTOR, size = N)` is a function that samples `size =N` observations from the vector named by `x=DATA.VECTOR`. 
- `replace = TRUE` : observations that are selected are returned to the vector and can be selected again
- `prob =` : a vector of the same length as the object being sampled `DATA.VECTOR` that gives the probability (0-1) of selecting each item.

```{r}
DATA.VECTOR <- seq(1,8)
sample(x= DATA.VECTOR, size = 8, replace = TRUE, prob = DATA.VECTOR/sum(DATA.VECTOR))
```

# `seq`

- Syntax: `seq(A,B,by=C)`
- Creates a sequence starting at `A` and ending at `B` incrementing by `C`.  Default is that `by=1`.

```{r}
seq(2,10,by=2)
seq(-5,-1)
```


# `tibble`
- Syntax `tibble(OBJECT)`
- Converts an object (usually a vector?) to a tibble for use with tidyverse.

```{r}
dat <- rbinom(n=1000, size=25, prob=.75)
is_tibble(dat)
dat <- tibble(dat)
is_tibble(dat)
```

# `unique`
- Syntax: `unique(DATA.VECTOR)` 
- Returns the set of unique values from the vector `DATA.VECTOR`.  It is a vector of length equal to the number of unique values.

```{r}
DATA.VECTOR <- c(1,2,3,4,3)
DATA.VECTOR
unique(DATA.VECTOR)
```

# `sample_n`
- Syntax: `sample_n(TIBBLE, size = SAMPLESIZE, replace = FALSE)`
- This will draw a random sample from the tibble `TIBBLE` of size `SAMPLESIZE` without replacement unless `replace=TRUE`.
- NOTE: that if we run `sample_n` in a "pipe" then we do not need to explicitly name `TIBBLE` as the pipe will identify its' relationship.

```{r}
sample.data1 <- sample_n(Pres2020.PV, size = 100, replace = TRUE)

sample.data2 <- Pres2020.PV %>%
               sample_n(size = 100, replace = TRUE)
```

# `mutate`
- Syntax: `mutate(TIBBLE, NEWVAR1 = MANIPULATION1, NEWVAR2 = MANIPULATION2)
- Updates or creates new variables/columns in a tibble in a specified way
- NOTE: that if we run `sample_n` in a "pipe" then we do not need to explicitly name `TIBBLE` as the pipe will identify its' relationship.
- NOTE: assigning the `TIBBLE` being mutated to the same `TIBBLE` will update/replace `TIBBLE`.  You could also create a new tibble.

```{r}
Pres2020.PV <- mutate(Pres2020.PV, 
                      margin = Biden - Trump,
                      AbsBidenError = abs(Biden - DemCertVote))

Pres2020.PV.new <- Pres2020.PV %>% 
                      mutate(
                        isOnlinePoll = ifelse(Mode == "Online",1,0),
                        AbsTrumpError = abs(Trump - RepCertVote))
```


# `summarize`
- Syntax: `summarize(TIBBLE, OBJNAME1 = FUNCTION1(VARIABLE1), OBJNAME1 = FUNCTION1(VARIABLE1))`
- Applies a function to a variable in `TIBBLE` and produces the object being assigned that function.
- NOTE: that if we run `summarize` in a "pipe" then we do not need to explicitly name `TIBBLE` as the pipe will identify its' relationship.
- NOTE: we can define the summarize as a new object to save it.

```{r}
summarize(Pres2020.PV, 
              AvgBiden = mean(Biden, na.rm=TRUE),
              BidenPct05 = quantile(Biden,.05),
              PctBidenLead = mean(Biden > Trump))

PollSummary <- Pres2020.PV %>%
                summarize(AvgBiden = mean(Biden, na.rm=TRUE),
                    BidenPct05 = quantile(Biden,.05),
                    PctBidenLead = mean(Biden > Trump))
```

# `rbinom`
- Syntax: `rbinom(n=N, size=SIZE, prob=PROB)
- Draws `N` samples of size `SIZE` from the binomial distribution where the probability of a 1 (Success) is given by `PROB`.  
- `PROB` can be either a value if the every draw in the `SIZE` has the same probability of occurring (e.g. flipping the same coin `SIZE` times), or it can vary in which case `PROB` is a vector of length `SIZE`
- NOTE: The output of this function is the number of 1's that are produced. To convert to a proportion you need to divide by `SIZE`.

```{r}
rbinom(n=10, size=100, prob=.75)
rbinom(n=10, size=100, prob=.75)/100
```

# `sum`
- Syntax: `sum(DATA.VECTOR)`
- Computes the sum of the numeric variable `DATA.VECTOR`
```{r}
DATA.VECTOR <- seq(1,8)
sum(DATA.VECTOR)
```

# `quantile`
- Syntax `quantile(VARIABLE, probs=PROBS)`
- For the variable named `VARIABLE` it computes the empirical quantiles of the variable given in the value (or vector!) `PROBS`

```{r}
quantile(Pres2020.PV$Biden, probs=.5) # calculate the 50th percentile (Median)
quantile(Pres2020.PV$Biden, probs=c(.025,.975)) # calculate the points containing 95% of the values
quantile(Pres2020.PV$Biden, probs=seq(0,1,by=.10)) # calculate the deciles
```




