---
title: "Resampling Unit"
author: "YOUR NAME HERE"
date: "9/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```

## Two Amazing Statistical Results are



# Two ways to sample

Whether we replace or not is important becasue....

```{r}
Z <- seq(1,8)
set.seed(42)
## Randomly draw 8 samples from Z, with replacement
sample(Z, 8, replace = TRUE) 
## Randomly draw 8 samples from Z, without replacement
sample(Z, 8, replace = FALSE) 
## Randomly draw 8 samples from Z, with replacement
sample(Z, 8, replace = TRUE, prob = Z/sum(Z))
``` 


# Birthday Problem: Solving via Simulation

What does each line do?  

To get larger class sizes what do you change?

To get more simulations what do you need to change?

Why does this look different than the one we saw in class?

Advanced: If you wanted to change the probability of sampling each day, what would you change?


```{r}
sims <- 10 ## number of sims
bday <- 1:365 ## possible bdays
answer <- NULL ## placeholder

for (k in 1:25) {
  count <- 0  ## counter
  for (i in 1:sims) {
    class <- sample(bday, k, replace = TRUE) 
    
    if (length(unique(class)) < length(class)) {
      count <- count + 1
    }}  
  ## printing the estimate
  cat("The estimated probability for", k,"people is:", 
      count/sims, "\n")
  answer[k] <- count/sims # store the answers
}
```


# Birthday Problem: Visualize

Why do we need to `bind_cols`?  What does that do?

```{r}
dat <- bind_cols(npeople=seq(1,25),answer)

ggplot(dat,aes(x=npeople,y=answer)) + 
  geom_line() +
  labs(x = "Number of People") +
  labs(y="Probability") + 
  labs(title = "Prob. of shared bday for each size") +
  geom_hline(yintercept=.5)
```


# Pennsylvania Voter File
```{r}
load("data/pa.sample.select.Rdata")
glimpse(pa.sample)
```

What does this do?  How differ from `transmute`?


```{r}
pa.sample <- pa.sample %>%
    mutate(likely.dem = ifelse(likely.party == "D",1,0), 
          likely.rep = ifelse(likely.party == "R",1,0), 
          likely.ind = ifelse(likely.party == "I",1,0))
```

Partisan composition in voter file

```{r}
PA.pty.breakdown <- pa.sample %>%
          summarize(pct.dem = mean(likely.dem), 
                    pct.rep = mean(likely.rep), 
                    pct.ind = mean(likely.ind))

PA.pty.breakdown
```

What is the difference between `sample` and `sample_n`?

What does this do?

How does `rand.dat` change over iterations?

If I wanted larger samples, how could you make that happen?


```{r}
samplesize <- seq(1,10)

for(i in seq_along(samplesize)){
  rand.dat <- sample_n(pa.sample, i , replace= TRUE)
}
dim(rand.dat)
```

Samples in a loop?  What am I looping over now?

What does `samplesize` do?

```{r}
dem.pty.est <- NULL
samplesize <- c(3,10,34,567,4762)

for(i in seq_along(samplesize)){
  rand.dat <- sample_n(pa.sample,
                samplesize[i],replace= TRUE)
  dem.pty.est[i] <- mean(rand.dat$likely.dem)
}
```

What does this show?

```{r}
samplesize <- seq(1,1000)
for(i in seq_along(samplesize)){
  rand.dat <- sample_n(pa.sample,i,replace= TRUE)
  dem.pty.est[i] <- mean(rand.dat$likely.dem)
}

dat <- bind_cols(samplesize,dem.pty.est)

pasample.plot <- dat %>% ggplot(aes(x=samplesize,y=dem.pty.est)) +
  geom_line() + 
  geom_point() +
  labs(x = "Number of Observations") + 
  labs(y = "Pct. Dem in Sample") 
```


# Thinking about error

```{r}
dat <- dat %>% 
  mutate(Absolute.Error = 
           abs(dem.pty.est - PA.pty.breakdown[[1]]),
         Squared.Error = 
           (dem.pty.est - PA.pty.breakdown[[1]])^2,
         cut = cut(samplesize,breaks = 10))

table(dat$cut)

dat %>% 
  group_by(cut) %>%
  summarize(avgae = mean(Absolute.Error),
            sdae = sd(Absolute.Error))
```

What does this do? Why are we doing this?

```{r}
# Fix Sample Size, Vary number of Samples
n.samplesize <- 1000
n.samples <- 5

est.sample5 <- NULL
for(i in 1:5){
    rand.dat <- sample_n(pa.sample,n.samplesize,replace= TRUE)
    est.sample5[i] <- mean(rand.dat$likely.dem)
}

est.sample5 <- tibble(est.sample5)
```

What does this show?

```{r}
est.sample5 %>% ggplot(aes(x=est.sample5)) + 
  geom_histogram(bins=5) +
  labs(title="Means for 5 samples") +
  xlim(.42,.53)
```

What does this do?

```{r}
est.sample5000 <- NULL
for(i in 1:5000){
  rand.dat <- sample_n(pa.sample,n.samplesize,replace= TRUE)
  est.sample5000[i] <- mean(rand.dat$likely.dem)
}

est.sample5000 <- tibble(est.sample5000)
```

# Central Limit Theorem

What is the Central Limit Theorem?

Why is it important?

```{r}
est.sample5000 %>% ggplot(aes(x=est.sample5000)) + 
  geom_histogram(bins=50) +
  labs(title="Means for 5000 samples") +
  xlim(.42,.53)
```

# Margin of Error

What does this do?

```{r}
n.samplesize <- 1000
my.sample <- sample_n(pa.sample,n.samplesize,
                      replace= TRUE)

B <- 5000
resample5000 <- NULL
for(i in 1:B){
  rand.dat <- sample_n(my.sample,n.samplesize,
                       replace= TRUE)
  resample5000[i] <- mean(rand.dat$likely.dem)
}
```

What does this do? What does it mean?

```{r}
quantile(resample5000,c(.25,.75)) 
quantile(resample5000,c(.05,.95)) 
quantile(resample5000,c(.025,.975)) 

```

What does this do? What does it mean?

```{r}
quantile(resample5000,c(.025,.975)) - mean(my.sample$likely.dem)
```


How many randomly sampled data points do you need to get a good estimate?

```{r}
samplesizes <- c(10,100,200,300,400,500,600,700,800,900,1000,5000,10000)
moe <- rep(NA,length(samplesizes))

for(i in seq_along(samplesizes)){
  my.sample <- sample_n(pa.sample,samplesizes[i],
                        replace= TRUE)
  
  resample1000 <- rep(NA,times=1000)
  for(j in 1:1000){
    rand.dat <- sample_n(my.sample,samplesizes[i],
                         replace= TRUE)
    resample1000[j] <- mean(rand.dat$likely.dem)
  }
  moe[i] <- quantile(resample1000,c(.975)) - mean(my.sample$likely.dem)
}
```

What does this do?  What does it mean?


```{r}
dat <- bind_cols(samplesizes,moe)

ggplot(dat) +
  geom_point(aes(x=samplesizes,y=moe)) +
  labs(x="Sample Size") + 
  labs(title = "Relationship Between Random Sample Size and Margin of Error") +
  scale_y_continuous(breaks=seq(0,.3,by=.02)) +
  labs(y="% Margin of Error (+/-)")
```