---
title: "Topic 6. Conditional Visualization"
author: "YOUR NAME HERE"
date: "9/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


NOTE: When you are done and have worked thru the chunks (and added notes), "knit" the document to create your own set of custom notes. Cool stuff!

Load the data and define some date objects.  Note the "month"/"date"/"year" formatting.
```{r}
load(file="data/Pres2020.PV.Rdata")
election.day <- as.Date("11/3/2020", "%m/%d/%Y")   
election.day16 <- as.Date("11/8/2016", "%m/%d/%Y") 
```

Now some wrangling...

```{r}
Pres2020.PV <- Pres2020.PV %>%
                mutate(EndDate = as.Date(Pres2020.PV$EndDate, "%m/%d/%Y"), 
                      StartDate = as.Date(Pres2020.PV$StartDate, "%m/%d/%Y"),
                      DaysToED = as.numeric(election.day - EndDate),
                      margin = Biden - Trump)
```

What type of plot is this?  Used for what kind of data?

```{r}
p <- ggplot(data = Pres2020.PV, aes(x = DaysToED)) +
  labs(title = "Number of 2020 National Polls Over Time") + 
  labs(x = "Number of Days until Election Day") + 
  labs(y = "Count") + 
  geom_bar(color="PURPLE") 
```


Can you fix the scale?

```{r}
p <- ggplot(data = Pres2020.PV, aes(x = DaysToED)) +
  labs(title = "Number of 2020 National Polls Over Time") + 
  labs(x = "Number of Days until Election Day") + 
  labs(y = "Count") + 
  geom_bar(color="PURPLE") 
```

Can you fix and flip the scale?
```{r}

```

What kind of graph is this? What is it used for?  What if I change `bins`?  Which, if any, is a good choice?

```{r}
p <- ggplot(data = Pres2020.PV, aes(x = DaysToED)) +
  labs(title = "Number of 2020 National Polls Over Time") + 
  labs(x = "Number of Days until Election Day") + 
  labs(y = "Count") + 
  geom_histogram(color="PURPLE",bins = 30) +
  scale_x_reverse()
```

#   Bivariate/Multivariate relationships

What is X? Y?

What type of graph do we use if continuous?  How would I change the variables being plotted here?  What does the following code do?

```{r}
margin.plot <- ggplot(Pres2020.PV, 
                      aes(x = EndDate, y = margin)) + 
  labs(title="Margin in 2020 Nat. Popular Vote Polls Over Time") + 
  labs(y = "Margin: Biden - Trump") + 
  labs(x = "Poll Ending Date")
```

How do I add points?
```{r}

```

What does this do?  Note that I have set `eval=FALSE` so you need to run that to get it to run!

```{r, eval = FALSE}
margin.plot  + 
    geom_point(color = "PURPLE")  + 
    scale_y_continuous(breaks=seq(-10,20,by=5)) +
    scale_x_date(date_breaks = "2 week", date_labels = "%b %d") 
```


How do I `jitter`? 

```{r}

```

Add `alpha` transparency?  
```{r}

```

More importantly, _why_ would I want to do this?

How do I add a line?  Do I want to?

```{r}

```

# Plotting multiple variables on the same plot - Time Series


```{r}
BidenTrumpplot <- ggplot(Pres2020.PV)  +
  labs(title="% Biden and Trump in 2020 National Popular Vote Polls Over Time") + 
  labs(y = "Pct. Support") + 
  labs(x = "Poll Ending Date") 
```

Now add points.  How come I add the x axis when I add the points?

```{r}
BidenTrumpplot <- BidenTrumpplot + 
  geom_point(aes(x = EndDate, y = Trump), 
             color = "red", alpha=.4)  +
  scale_x_date(date_breaks = "2 week", date_labels = "%b %d")
```

Now add `Biden` points.  Do I need to add the axis again?  Why or why not?


```{r}

```


Change the y-axis to range from (0,100)?  How about (0,200)?  What is a good choice for this range? _Why_?

```{r}

```

Make your best possible visualization.  Think about what you are doing and why you are doing it!  Can you save it as a PDF?  (See last class!)

```{r}

```


# Smoothing & Conditional Means

What is smoothing?  Why do we want to do it?  What is the basic idea?


# Looping

Very bad code.  Why?

```{r,eval=FALSE}
mean(Pres2020.PV$Trump[Pres2020.PV$EndDate == "2020-10-31"])
mean(Pres2020.PV$Trump[Pres2020.PV$EndDate == "2020-11-01"])
mean(Pres2020.PV$Trump[Pres2020.PV$EndDate == "2020-11-02"])
```

Saving you from typing mistakes. Use this in the baby loops we are doing....

```{r}
dates <- c("2020-10-31","2020-11-01","2020-11-02")
```

What are the elements of a loop?

Can you program a loop?
```{r}

```

# Looping for real: Creating a polling smoother

Saving you from typing mistakes.  Can you follow every line of code?   

What is `Bandwidth`?
What is `PV_avg`?

What is the loop doing?

```{r}
all_dates <- seq(min(Pres2020.PV$EndDate), election.day, 
                 by = "days")
Bandwidth <- 2
PV_avg <- vector(length(all_dates), mode = "list") # holding variable

for (i in seq_along(all_dates)) {
  date <- all_dates[i]

    PV_avg[[i]] <- Pres2020.PV %>%
    filter(as.integer(EndDate - date) <= 0 &
           as.integer(EndDate - date) > - Bandwidth) %>%
    summarize(Biden = mean(Biden),
              Trump = mean(Trump)) %>%
      mutate(date = date)
}
```


What is this doing? Why is this importnat?
```{r}
pop_vote_avg <- bind_rows(PV_avg)
```

Are you ready to plot?  What does this do?  What does each line do?  Can you save this as a PDF?

```{r}
PlotTS <- ggplot(pop_vote_avg) + 
  geom_line(aes(x = date, y = Trump), color = "red") + 
  geom_line(aes(x = date, y = Biden), color = "blue")  + 
  labs(title="3-Day Avg. Support for Biden and Trump in 2020 National Popular Vote Polls Over Time") + 
  labs(y = "Avg. Support") + 
  labs(x = "Poll Window Ending Date (3 Days)") + 
  scale_x_date(date_breaks = "2 week", date_labels = "%b %d") + 
  scale_y_continuous(breaks=seq(30,70,by=5))
```


How would you change the scale of y to go from 0 to 100? 
How would you change the color of Biden to "GREEN" and Trump to "YELLOW"?
How would you change the frequency of the x-axis labels to be every 3 weeks?

# Blow the Doors Off!

Change the bandwidth to the one you think is best!  Why did you make that choice?  Was it data-driven or based on your understanding of elections and campaigns in the US?

What is the best visualization you can make using the tools we covered?  You can do it!
