---
title: "Workflows"
author: "Will Doyle"
date: "10/28/2021"
output: html_document
---

This guide will provide the quickest "through line" when using workflows.

```{r}
library(tidyverse)
library(tidymodels)
```

```{r}
mv_df<-read_rds("mv.Rds")%>%
  filter(!is.na(budget))%>%
  mutate(log_gross=log(gross))
```

## Example 1: Simple linear regression

First, we'll split the data into training and testing sets:

```{r}
split_data<-initial_split(mv_df)

mv_train<-training(split_data)

mv_test<-testing(split_data)
```


## Define Model

Next, let's define the model we want to use: OLS regression

```{r}
lm_fit <- 
  linear_reg() %>% 
  set_engine("lm")%>%
  set_mode("regression")
```

## Define Recipe

A recipe is a set of steps that gets the data ready for analysis. A recipe requires a formula that defines the outcomes and the predictors, and a dataset. We'll do two: take the log of budget, and create a series of dummy variables for rating. 
```{r}
mv_formula<-as.formula("log_gross~budget+rating")

mv_recipe<-recipe(mv_formula,mv_train)%>%
  step_log(budget)%>%
  step_dummy(rating)

```


## Create workflow

A workflow contains all of these steps, allowing us to put them together in a unified framework. Our workflow will start with the model and the recipe.

```{r}
movie_wf<-workflow()%>%
  add_model(lm_fit)%>%
  add_recipe(mv_recipe)
```

## Fit workflow to training data

Now, we can run this model on our training data,
establishing the relationship between our predictors (budget and rating) and the outcome (log of gross).

```{r}
movie_wf<-movie_wf%>%
  fit(mv_train)
```

## Check model fit in testing data

We want to check our predictions against the testing dataset. Since we have the full testing dataset in hand, we can do it using last_fit. 
```{r}
movie_lf<-last_fit(movie_wf,split_data)
movie_lf$.metrics
```

For your exercise, you'll need to use the testing dataset we gave you, which will work like this:
```{r}
mv_test<-
  movie_wf%>%
  predict(new_data=mv_test)%>%
  bind_cols(mv_test)

mv_test
```

The `mv_test` file now contains a new prediction for ever case in the testing dataset. 